
<!-- This .Rmd file is set up to run either independently by checking for 
the presence of necessary objects in the global environment and running 
related scripts if not, or also as a child to the "main-doc". For that latter
reason, comments and headings that are not relevant to a final report to
pubic audiences are suppressed or moved to code chunks that can optionally be 
`echo`ed as desired. -->

```{r}
# set flags for running this script
print_documentation <- FALSE
```

```{r}
# run scripts for necessary packages and objects as necessary
if (!"chHexs"        %in% objects()) source("settings--main.R", echo = FALSE)
if (!"my_state_abbr" %in% objects()) source("settings--profile.R", echo = FALSE)
if (!"bin_age"       %in% objects()) source("method--general-helper-functions.R", echo = FALSE)

Sys.setenv(census_api_key = census_key)
# Check to see that the expected key is output in your R console
Sys.getenv("census_api_key")

options(scipen=999)

```

```{r identify relevant fips codes for population pulls}
my_state_fip <- 
  fips_codes %>% 
  filter(state == my_state_abbr) %>% 
  pull(state_code) %>% 
  unique()

if ("my_county" %in% objects()) {
  my_county_fip <- 
    fips_codes %>% 
    filter(state == my_state_abbr, 
           str_detect(county, my_county)) %>% 
    pull(county_code) %>% 
    unique() 
}

if (exists("my_county_fip")) {
  county_arg <- as.numeric(my_county_fip)
} else {
  county_arg <- NULL
}

```

The following code compares block-to-tract aggregated values from the 2020 DHC and tract-level values for the 2020 DHC to ensure there are no discrepancies.

```{r compare block-to-tract aggregated values and tract-level values for 2020 DHC}

#r pull single-age population counts aged 0, ..., 17 from 2020 DHC
#Load variables for 2020 DHC
#Also available at https://api.census.gov/data/2020/dec/dhc/variables.html
var_dhc_2020 <- load_variables(2020, dataset = "dhc")

#Filter variables
var_dhc_2020 <- 
  var_dhc_2020 %>% 
  filter(str_detect(name, "P14")) %>% 
  mutate(age = str_extract(label, "\\d+"),
         age = as.numeric(age) - str_detect(label, "Under"),
         sex = case_when(str_detect(label, "Female") ~ "female",
                         str_detect(label, "Male")   ~ "male",
                         TRUE                        ~ "all")) %>% 
  filter(age %in% 0:17,
         !(sex == "all"))

#Block-level pop count by age
pop_block_2020_dhc <-
  get_decennial(
    geography = "block",
    variables = var_dhc_2020$name,
    year = 2020,
    sumfile = "dhc",
    state  = my_state_fip,
    county = county_arg,
    geometry = F) %>% 
  merge(var_dhc_2020 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>% 
  rename(GEOID_BL20 = GEOID)

#Tract-level pop count by age
pop_tract_2020_dhc <-
  get_decennial(
    geography = "tract",
    variables = var_dhc_2020$name,
    year = 2020,
    sumfile = "dhc",
    state  = my_state_fip,
    county = county_arg,
    geometry = F) %>% 
  merge(var_dhc_2020 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>% 
  rename(GEOID_TR20 = GEOID)
  
# Table ID starting with `P`   is available at the as low as block-level
# Table ID starting with `PCT` is available at the as low as tract-level
# P14:   Count by single-year age for the population under 20 years (available at the block/tract level)
# PCT12: Count by single-year age for the total population (available at the tract level); values are equivalent to P14 at the tract level

#aggregate block-level 2020 DHC data to tract-level 
pop_block_2020_dhc_to_tract <-
  pop_block_2020_dhc %>%
  mutate(GEOID_TR20 = as.numeric(substr(GEOID_BL20, 1, 11))) %>%
  group_by(GEOID_TR20, age, sex) %>%
  summarise(value = sum(value))

#merge two dataset
comp <- 
  merge(pop_block_2020_dhc_to_tract,
        pop_tract_2020_dhc,
        by = c("GEOID_TR20", "age", "sex"),
        suffixes = c("_bltotr", "_tract")) %>%
  select(-variable, -NAME)

#check if block_to_tract and tract values are equivalent
comp %>%
  group_by(GEOID_TR20, age, sex) %>%
  summarise(equiv = all(value_bltotr == value_tract)) %>%
  tabyl(equiv)

```

The following codes are generating population counts of certain age groups with 2020 Census PL Redistricting Data and ACS 5-year. This used to be a temporary solution before the 2020 Census DHC was released. As the 2020 Census DHC was released, we compare the population counts from the DHC and the estimates of RD+ACS5.
* 2020 Census PL Redistricting Data: to get tract-level population of kids aged 0-17
* 2018-2022 ACS5: to get age group break-downs by tract
* Combine the above two to get tract-level population of age groups 0-2, 3-4, 5, 6-11, 12-17

```{r pull tract-level 0-17 population data from 2020 PL 94-171 Redistricting Data}
#Load variables for 2020 PL RD
#Also available at https://api.census.gov/data/2020/dec/pl/variables.html
var_rd_2020 <- load_variables(2020, dataset = "pl")

var_rd_2020 <-
  var_rd_2020 %>%
  filter(name %in% c("P1_001N", # Table P1 reflects RACE, and field P1_001N represents Total
                     "P3_001N"  # Table P3 reflects RACE FOR THE POPULATION 18 YEARS AND OVER, and field P3_001N represents Total
                     ))

pop_tract_2020_rd <- 
  get_decennial(
    geography = "tract",
    variables = var_rd_2020$name,
    year = 2020,
    sumfile = "pl",
    state = my_state_fip,
    county = county_arg,
    geometry = FALSE) %>%
  merge(var_rd_2020 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>% 
  rename(GEOID_TR20 = GEOID)

```

```{r rename 2020 field and calculate lt 18 overall population, echo = FALSE}
pop_tract_2020_rd <- 
  pop_tract_2020_rd %>%
  mutate(age = ifelse(str_extract(variable, "^.{2}") == "P1", "Total Population",
               ifelse(str_extract(variable, "^.{2}") == "P3", "18 Plus", NA))) %>%
  select(-variable, -NAME) %>%
  pivot_wider(names_from = age,
              values_from = value) %>% 
  mutate(totallt18_2020 = `Total Population`-`18 Plus`)

# Examine structure
if (FALSE) {
  head(pop_tract_2020_rd)
}

```

```{r compare tract-level data from PL RD and DHC}
#aggregate single-year age tract-level pop counts
pop_tract_2020_dhc_test <-
  pop_tract_2020_dhc %>%
  group_by(GEOID_TR20) %>%
  summarise(totallt18_2020 = sum(value))

comp <- 
  merge(pop_tract_2020_rd, 
        pop_tract_2020_dhc_test, 
        by = "GEOID_TR20",
        suffixes = c("_redis", "_dhc"))

# Examine the differences
summary(comp$totallt18_2020_redis)
summary(comp$totallt18_2020_dhc)

ggplot(comp,
       aes(x = totallt18_2020_redis,
           y = totallt18_2020_dhc)) +
  geom_point(alpha = 0.3) 

# confirmed the 0-17 aggregate of pop counts of DHC is equivalent to pop counts(0-17) of the PL RD
rm(pop_tract_2020_dhc_test, comp)

```

```{r pull ACS 5-year data on age breakdowns}
#Load variables for most recent ACS5
#Also available at https://api.census.gov/data/2022/acs/acs5/variables.html
var_acs5 <- load_variables(acs5_year, dataset = "acs5")

#Filter variables
#Age groups are pulled from ACS 5-year table B09001 Population Under 18 Years by Age
var_acs5 <- 
  var_acs5 %>% 
  filter(str_detect(name, "B09001")) %>% 
  mutate(inhh = str_extract(label, "In households")) %>% 
  filter(!is.na(inhh))

pop_by_age_acs5 <- 
  get_acs(geography = "tract",
          variables = var_acs5$name,
          state = my_state_fip,
          county = county_arg,
          year = acs5_year) %>%
  merge(var_acs5 %>% select(-geography, -inhh),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>% 
  rename(GEOID_TR20 = GEOID)

```

```{r combine acs5 variables into custom age groups}
#Group ages into 0-2, 3-4, 5, 6-11, and 12-17.
pop_by_agegroup_acs5 <- 
  pop_by_age_acs5 %>%
  mutate(variable = recode(variable,
                           "B09001_002" = "total",
                           "B09001_003" = "pop_02",
                           "B09001_004" = "pop_34",
                           "B09001_005" = "pop_5",
                           "B09001_006" = "pop_611",
                           "B09001_007" = "pop_611",
                           "B09001_008" = "pop_1217",
                           "B09001_009" = "pop_1217")) %>%
  group_by(GEOID_TR20, variable) %>%
  summarise(estimate = sum(estimate),
            moe      = moe_sum(moe = moe,
                               estimate = estimate))

# Examine structure
head(pop_by_agegroup_acs5)  

#check
pop_by_agegroup_acs5 %>%
  ungroup() %>% 
  summarise(total  = sum(estimate[variable == "total"]),
            pop_02 = sum(estimate[variable == "pop_02"]),
            pop_34 = sum(estimate[variable == "pop_34"])) %>%
  mutate(percent_0004 = (pop_02 + pop_34)/total*100 %>% round(1))

```

```{r ACS age as percent of total, echo = FALSE, message = FALSE}
#For each age group find their percent of the total

# /!\ Could arguably do this in long form, with a `group_by(GEOID)`, and dispensing
# with the `total` row, replacing it with sum(estimate)
pop_agegroup_pcts_acs5 <-
  pop_by_agegroup_acs5 %>%
  pivot_wider(names_from = variable,
              values_from = c(estimate, moe)) %>%
  mutate(per02_acs   = estimate_pop_02   / estimate_total,
         per34_acs   = estimate_pop_34   / estimate_total,
         per5_acs    = estimate_pop_5    / estimate_total,
         per611_acs  = estimate_pop_611  / estimate_total,
         per1217_acs = estimate_pop_1217 / estimate_total,

         # Reference for arguments -- moe_prop(num, denom, moe_num, moe_denom)         
         per02_moe   = moe_prop(estimate_pop_02,   estimate_total, moe_pop_02,   moe_total),
         per34_moe   = moe_prop(estimate_pop_34,   estimate_total, moe_pop_34,   moe_total),
         per5_moe    = moe_prop(estimate_pop_5,    estimate_total, moe_pop_5,    moe_total),
         per611_moe  = moe_prop(estimate_pop_611,  estimate_total, moe_pop_611,  moe_total),
         per1217_moe = moe_prop(estimate_pop_1217, estimate_total, moe_pop_1217, moe_total),
         
         #Here we are doing a crude cut of the data to get our desired age groups of 6-12 and 13-17
         cut_1217          = estimate_pop_1217 / 6,
         
         # Note for apportioning moe, that the total noise is proportional to 
         # variances, and not moe --
         #   var(a1 + a2 + ... + aN) 
         #     = var(a1) + var(a2) + ... var(aN) 
         #     = N*var(a1) if variances are assumed the same (and correlation = 0)
         # Thus, we need to convert MOE calculations to variances to apportion,
         # and then convert them back
         per1217_se        = per1217_moe / qnorm(0.95), # note: qnorm(0.95) ~== 1.645
         per1217_var       = per1217_se^2,
         cut_1217_var      = per1217_var / 6,
         cut_1217_moe      = sqrt(cut_1217_var) * qnorm(0.95),
         
         estimate_pop_612  = estimate_pop_611  + cut_1217,
         estimate_pop_1317 = estimate_pop_1217 - cut_1217,
         moe_pop_612       = per611_moe        + cut_1217_moe,
         moe_pop_1317      = per1217_moe       - cut_1217_moe,
         
         per612_acs  = estimate_pop_612  / estimate_total,
         per1317_acs = estimate_pop_1317 / estimate_total,
         per612_moe  = moe_prop(estimate_pop_612,  estimate_total, moe_pop_612,  moe_total),
         per1317_moe = moe_prop(estimate_pop_1317, estimate_total, moe_pop_1317, moe_total)) %>%
  select(GEOID_TR20, per02_acs, per34_acs, per5_acs, per612_acs, per1317_acs,
                     per02_moe, per34_moe, per5_moe, per612_moe, per1317_moe)

```

```{r address not-a-number values due to division, echo = FALSE, message = FALSE}
sapply(pop_agegroup_pcts_acs5, function(x) mean(is.nan(x))) %>% sort()
#Get replace a few NaNs with zeros. No kids here (tracts:980000,980100,381700)

pop_agegroup_pcts_acs5 <- replace(pop_agegroup_pcts_acs5, is.na(pop_agegroup_pcts_acs5), 0)

```

```{r add acs to 2020 pop, echo = FALSE, message = FALSE}
#Merge the 2020 population and ACS age group percents
pop_2020_rd_acs5 <-
  merge(x = pop_tract_2020_rd,
        y = pop_agegroup_pcts_acs5,
        by = "GEOID_TR20",
        all.x = TRUE)

# Examine the structure
head(pop_2020_rd_acs5)

```

```{r estimate 2020 age breakdowns, echo = FALSE, message = FALSE}
#Use ACS percents to split the 2020 PL RD less than 18 population into age groups

pop_age_groups_2020_rd_acs5 <-
  pop_2020_rd_acs5 %>%
  mutate(est_2020_02   = round(totallt18_2020*per02_acs,   0),
         est_2020_34   = round(totallt18_2020*per34_acs,   0),
         est_2020_5    = round(totallt18_2020*per5_acs,    0),
         est_2020_612  = round(totallt18_2020*per612_acs,  0),
         est_2020_1317 = round(totallt18_2020*per1317_acs, 0),
         
         # Note: reference for arguments is `moe_product(est1, est2, moe1, moe2)`
         # This implicitly assumes that 
         moe_2020_02   = moe_product(totallt18_2020, per02_acs,   0, per02_moe),
         moe_2020_34   = moe_product(totallt18_2020, per34_acs,   0, per34_moe),
         moe_2020_5    = moe_product(totallt18_2020, per5_acs,    0, per5_moe),
         moe_2020_612  = moe_product(totallt18_2020, per612_acs,  0, per612_moe),
         moe_2020_1317 = moe_product(totallt18_2020, per1317_acs, 0, per1317_moe)
         ) %>%
  select(GEOID_TR20,
         est_2020_02, est_2020_34, est_2020_5, est_2020_612, est_2020_1317,
         moe_2020_02, moe_2020_34, moe_2020_5, moe_2020_612, moe_2020_1317) %>%
  pivot_longer(cols = -GEOID_TR20,
               names_to  = "variable",
               values_to = "value") %>%
  separate(variable, 
           into = c("variable", "year", "agegroup")) %>%
  pivot_wider(names_from  =  c("variable","year"),
              values_from = "value",
              names_sep   =  "_") %>%
  mutate(
    agegroup_code = recode(agegroup,
                           "02"   = 1,
                           "34"   = 2 ,
                           "5"    = 3,
                           "612"  = 4,
                           "1317" = 5))

```

```{r compare pop est by age groups from ACS5+RD vs. DHC, echo = FALSE, message = FALSE}
pop_age_groups_2020_dhc <-
  pop_tract_2020_dhc %>% 
  select(-variable, NAME) %>%
  mutate(agegroup = 
           case_when(between(age,  0,  2) ~ "02",
                     between(age,  3,  4) ~ "34",
                     age == 5             ~ "5",
                     between(age,  6, 12) ~ "612",
                     between(age, 13, 17) ~ "1317")) %>% 
  group_by(GEOID_TR20, agegroup) %>%
  summarise(value = sum(value))
  
comp_dhc_redis_acs5 <- 
  merge(pop_age_groups_2020_dhc,
        pop_age_groups_2020_rd_acs5 %>% rename(value = est_2020),
        by = c("GEOID_TR20", "agegroup"),
        suffixes = c("_dhc", "_redis_acs5"))

ggplot(comp_dhc_redis_acs5,
       aes(x = value_redis_acs5,
           y = value_dhc)) +
  geom_abline() + 
  geom_point(alpha = 0.3) +
  geom_smooth()

rm(comp_dhc_redis_acs5)
rm(var_rd_2020, pop_tract_2020_rd, var_acs5, pop_by_age_acs5, pop_by_agegroup_acs5, pop_agegroup_pcts_acs5, pop_2020_rd_acs5)

```

The following code compares block-to-tract aggregated values from the 2010 SF1 and tract-level values for the 2010 SF1 to ensure there are no discrepancies.

```{r compare block-to-tract aggregated values and tract-level values for 2010 SF1}
#r pull single-age population counts aged 0, ..., 17 from 2010 SF1

#Load variables for 2020 DHC
var_sf1_2010 <- load_variables(2010, dataset = "sf1")

#Filter variables
var_sf1_2010 <- 
  var_sf1_2010 %>% 
  filter(str_detect(name, "P014")) %>% #name, ("PCT012|P012|P014")
  mutate(age = str_extract(label, "\\d+"),
         age = as.numeric(age) - str_detect(label, "Under"),
         age = ifelse(is.na(age), "all", age),
         sex = case_when(str_detect(label, "Female") ~ "female",
                         str_detect(label, "Male")   ~ "male",
                         TRUE                        ~ "all")) %>% 
  filter(age %in% 0:17,
         !(sex == "all"))

#Block-level pop count by age
pop_block_2010_sf1 <-
  get_decennial(
    geography = "block",
    variables = var_sf1_2010$name,
    year = 2010,
    sumfile = "sf1",
    state  = my_state_fip,
    county = county_arg,
    geometry = F) %>% 
  merge(var_sf1_2010 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>%
  rename(GEOID_BL10 = GEOID) %>% 
  mutate(GEOID_TR10 = str_sub(GEOID_BL10, 1, 11)) # get tract-level GEOID

#Block-group-level pop count by age
pop_bg_2010_sf1 <-
  get_decennial(
    geography = "block group",
    variables = var_sf1_2010$name,
    year = 2010,
    sumfile = "sf1",
    state  = my_state_fip,
    county = county_arg,
    geometry = F) %>% 
  merge(var_sf1_2010 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>% 
  rename(GEOID_BG10 = GEOID) %>%
  mutate(GEOID_TR10 = str_sub(GEOID_BG10, 1, 11)) # get tract-level GEOID

#Tract-level pop count by age
pop_tract_2010_sf1 <-
  get_decennial(
    geography = "tract",
    variables = var_sf1_2010$name,
    year = 2010,
    sumfile = "sf1",
    state  = my_state_fip,
    county = county_arg,
    geometry = F) %>% 
  merge(var_sf1_2010 %>% select(-label, -concept),
        by.x = "variable",
        by.y = "name",
        all = TRUE) %>%
  rename(GEOID_TR10 = GEOID)

#aggregate block-level 2010 SF1 data to tract-level 
pop_block_2010_sf1_to_tract <-
  pop_block_2010_sf1 %>%
  mutate(GEOID_TR10 = as.numeric(substr(GEOID_BL10, 1, 11))) %>%
  group_by(GEOID_TR10, age, sex) %>%
  summarise(value = sum(value))

#merge two dataset
comp <- 
  merge(pop_block_2010_sf1_to_tract,
        pop_tract_2010_sf1,
        by = c("GEOID_TR10", "age", "sex"),
        suffixes = c("_bltotr", "_tract")) %>%
  select(-variable, -NAME)

#check if block_to_tract and tract values are equivalent
comp %>%
  group_by(GEOID_TR10, age, sex) %>%
  summarise(equiv = all(value_bltotr == value_tract)) %>%
  tabyl(equiv)

rm(pop_block_2010_sf1_to_tract, comp)

```

```{r compare these population estimates for 4-year-olds to the estimating 4-year-olds by cutting 3+4 in half}
if (FALSE) {
  load(glue("{output_path}pop_data_{my_output_tag}.Rda"))
  pop_by_age_sep <- pop_by_age 
  rm(pop_by_age)
  load(glue("{output_path}pop_data_{my_output_tag}_20240227.Rda"))
  comp_pop <- 
    pop_by_age %>% select(GEOID_TR20, age_4_count) %>% 
    merge(pop_by_age_sep %>% select(GEOID_TR20, age_4_count),
          by = "GEOID_TR20",
          #all = TRUE,
          suffixes = c(".half", ".sep"))
  
  ggplot(comp_pop,
         aes(x = age_4_count.half,
             y = age_4_count.sep)) +
    geom_point(alpha = 0.3) + 
    geom_45(color = "blue") +
    geom_smooth(color = "red") +
    theme_minimal()
}

```
