# Background

This R script is designed to read all of the customized settings that a user has provided in a given Excel interface to ELPEP, and to set off programming code.

To run this script--and consequently ELPEP:

1. provide the name of Excel file that contains your customizations in the indicated area of the gray box below, 
2. make sure that you have taken all other steps to download necessary files described at https://www.github.com/chapinhall/elpep, and then
3. in the menu above go to "Code >> Run Region >> Run All"

```{r user specified-elpep customization file}
elpep_customization_file <- "C:/Users/hkang/Chapin Hall Center for Children/CCPRP Project - Documents/ELPEP/Illinois/Interface for Specifying Custom ELPEP Eligibility Estimates.xlsm"
```

```{r}
# These lines reset the run and load all packages necessary for further steps
rm(list = setdiff(ls(), "elpep_customization_file"))
source("settings--main.R")
```


```{r}
# This line runs a script to harvest all details specified in the customization file
source("method--read-customization-file.R")
```

```{r}
source("method--general-helper-functions.R")
rmarkdown::render(
  input = "run--00--main-doc.Rmd",
  output_file = glue("{output_path}ELPEP Technical Report -- {my_state} -- {my_output_tag}.html"))
```


