# Background

This R script is designed to read all of the settings customizations that a user has provided in a given Excel interface to ELPEP, and to set off programming code.

To run this script--and consequently ELPEP:

1. provide the name of Excel file that contains your customizations in the indicated area of the gray box below, 
2. make sure that you have taken all other steps to download necessary files described at https://www.github.com/chapinhall/elpep, and then
3. in the menu above go to "Code >> Run Region >> Run All"

```{r user specified-elpep customization file}
#elpep_customization_file <- "Interface for Specifying Custom ELPEP Eligibility Estimates.xlsm"
elpep_customization_file <- "C:/users/nmader/Downloads/Interface for Specifying Custom ELPEP Eligibility Estimates--Maine.xlsm"
```


# Technical Content

Load all packages necessary for ELPEP--

```{r}
source("settings--main.R")
```

Establish function for harvesting customizations from the Excel spec file--

```{r functions to parse contents from the excel file}
getRange <- function(file, range, istable = FALSE, header = TRUE) {
  if (istable) {
    val <- read.xlsx(file, namedRegion = range, colNames = header)
  } else {
    val <- read.xlsx(file, namedRegion = range, colNames = FALSE)[, 1]
  }
  return(val)
}
getInfo <- function(i, ...) getRange(file = elpep_customization_file, i, ...)

```


Read all customization parameters from the Excel spec file.

```{r}
elpep_spec_names <- getNamedRegions(elpep_customization_file)

elpep_params <- NULL
for (i in 1:length(elpep_spec_names)) {
  esn <- elpep_spec_names[i]
  istable <- 
    !str_detect(
      attr(elpep_spec_names, "position")[i], 
      "^\\w+\\d+$")
  
  suppressWarnings(assignment_try <- try(assign(esn, getInfo(esn, istable = istable))))
  if (is.null(assignment_try) || (length(assignment_try)==1 && !str_detect(assignment_try, "Error"))) {
    elpep_params <- c(elpep_params, esn)
  }
}
```

```{r set defaults as necessary}

```

```{r recast to correct data types}
inputs_to_numericize <- c("kid_age_thres_p")
for (i in inputs_to_numericize) {
  assign(i, as.numeric(get(i)))
}
```


```{r recode true and false}
inputs_tf <- c("use_only_sae_model_estimates",
               "developer_mode",
               "rerun_sae")

for (i in inputs_tf) {
  assign(i,
         case_match(get(i),
                    "True"  ~ TRUE,
                    "False" ~ FALSE,
                    TRUE    ~ NA))
}
```

```{r vectorize comma-separated inputs}
inputs_to_vectorize <- 
  c("local_ccdf_incratio_cuts",
    "output_fpl_cuts")

for (i in inputs_to_vectorize) {
  assign(i,
         str_split(get(i), pattern = ",") %>% trim() %>% as.numeric())
}
```


```{r prepare inputs for age aggregation}

```

```{r prepare inputs for map generation}
map_geos <- 
  map_geos %>% 
  str_split(pattern = ",") %>% 
  str_trim()

map_geo_level <-
  case_match(map_geo_level,
             "School District(s)"                      ~ "school",
             "Zip Code(s)"                             ~ "zcta",
             "County (or Counties)"                    ~ "county",
             "Custom Geography (to be provided below)" ~ "aux")



```



Kick off the run

```{r}
source("method--general-helper-functions.R")
rmarkdown::render(
  input = "run--00--main-doc.Rmd",
  output_file = glue("{output_path}ELPEP Technical Report -- {my_state} -- {my_output_tag}.html")
  )
```


