
<!-- This .Rmd file is set up to run either independently by checking for 
the presence of necessary objects in the global environment and running 
related scripts if not, or also as a child to the "main-doc". For that latter
reason, comments and headings that are not relevant to a final report to
pubic audiences are suppressed or moved to code chunks that can optionally be 
`echo`ed as desired. -->

```{r}
# set flags for running this 01b script
print_documentation <- FALSE
```

```{r }
# run scripts for necessary packages and objects as necessary
if (!"chHexs"        %in% objects()) source("settings--main.R", echo = FALSE)
if (!"my_state_abbr" %in% objects()) source("settings--profile.R", echo = FALSE)
if (!"bin_age"       %in% objects()) source("method--general-helper-functions.R", echo = FALSE)
```

```{r read in acs1 data and documentation}
my_pums <- 
  pums_variables %>% 
  filter(year == 2019, # base_year, ... currently using 2019 because 2021 documentation is not yet available for the current version of tidycensus
         survey == "acs1") %>% 
  select(var_code, var_label, level, val_min, val_max, val_label)

acs1_save_file <- glue("{input_path}census_acs1_{base_year}.csv")
if (!file.exists(acs1_save_file)) {
  source("pull--acs1-microdata.R")
} else {
  acs1 <- fread(file = acs1_save_file)
}
```

```{r ensure that ACS1 PUMA values are character to be consistent with other datasets}
acs1[j = PUMA := str_pad(PUMA, width = 5, side = "left", pad = "0")]
```


### Sample Restrictions

ACS 1-Year microdata for `r base_year` are downloaded from the [Census Application Program Interface (API)](https://www.census.gov/data/developers/data-sets/acs-1year.html) and subset to households not living in group quarters, and to `r my_state_abbr` records.

```{r subset microdata}
# TYPE (Census) -- 1 - Housing unit, 2 - group quarters
# Sample restricted to housing unit only
acs1 <- 
  acs1[TYPE %in% c(1) &
         ST == my_state_fip &
         YEAR == max(YEAR)]

# If county has been specified, load geographic data to perform subsetting to
# relevant PUMAs
if (exists("my_county")) {
  load(glue("{output_path}geo_data_{my_output_tag}.Rda"))
  acs1 <- acs1[PUMA %in% geo_crosswalk$PUMA]
}
```

### Calculation of Household-Level Measures

```{r view variables relevant to head-of-household coding, results = "asis", eval = developer_mode}
if (print_documentation) {
  my_pums %>% 
    filter(var_code %in% c("RELSHIPP", "SCHL", "INDP", "RAC2P", "HISP")) %>% 
    kable()
}
```

```{r investigate the number of households without a householder, eval = developer_mode}
acs1[by = .(YEAR, SERIALNO),
     j  = .(has_hoh = maxNA(RELSHIPP == 20))] %>% 
  summary()
```

Although not currently used in this analysis, this codebase develops indicators pertaining to the head of household, including education, race/ethnicity, and industry of employment (if applicable). In other future specifications, these may be considered for inclusion in measures predicted by the SAE method and, in turn, used to capture macroeconomic trends in the "now"casting step.
<!-- /*\ Note: this note about modeling is specific to the current implementation in Illinois, and may be of interest for modifying locally  -->

```{r prepare ACS 1-year microdata for head of household}
# Note: for households that contain multiple families, this method may produce inaccurate results (i.e., head of household may be unrelated to child in multi-family households).
acs1_hoh <- 
  acs1 %>%
  .[RELSHIPP == 20] %>%
  .[, `:=`(hoh_educ = 
             case_when(between(SCHL,  2, 14) ~ "less_hs", 
                       between(SCHL, 15, 17) ~ "hs", 
                       between(SCHL, 18, 20) ~ "somecoll", 
                       between(SCHL, 21, 21) ~ "coll", 
                       between(SCHL, 22, 24) ~ "collplus", 
                       TRUE ~ "NA"), # Default value is NA
           hoh_industry = bin_industry(INDP), # bin_industry() is defined in the `helper` script
           hoh_race =
             case_when(RAC2P == 1 ~ "white", 
                       RAC2P == 2 ~ "black", 
                       TRUE ~ "other_multiple"),
           hoh_hispanic = 1*(HISP != 1))] %>% # Note: multiplication by 1 converts the TRUE/FALSE output to 1/0
  .[j = hoh_industry_vulnerable := ifelse(hoh_industry == "NA", 
                                          "NA", 
                                          str_detect(hoh_industry, "educhealth|food"))] %>% # Note: the second argument to str_detect() uses regular experessions. For reference, see e.g. https://regexone.com/
  dplyr::select(YEAR, SAMPLE, SERIALNO, hoh_educ, hoh_industry, hoh_industry_vulnerable, hoh_race, hoh_hispanic)
```

```{r view variables relevant to constructing household-level measures, eval = developer_mode, results = "asis"}
if (print_documentation) {
  my_pums %>% 
    filter(var_code %in% c("RELSHIPP", "SCH", "ESR")) %>%
    kable()
}
```

An additional set of indicators is built using all information about household members to capture not just family income (which the ACS has in a field reported even for younger members of the household) but also presence and labor force status of adults in the household:

* Presence of spouse is determined based on either an opposite- or same-sex spouse to the head of household
* Qualification of head of household or spouse in work or school is determined by either:
  - civilian employment either "at work" or "with a job but not at work"
  - enrollment in public or private school or college

<!-- /!\\ Note that the Chapin Hall research team is currently working on separating income and family relationships information for "subfamilies" that may be residing together with the primary family which the Head-of-Household that is the main respondent belongs to. -->

```{r prepare ACS 1-year microdata for families}
# Key field values (Census)
# ESR -- 0 - N/A (less than 16 yrs old), 1 - Civilian employed, at work, 2 - Civilian employed, with a job but not at work, 3 - Unemployed, 4 - Armed forces, at work, 5 - Armed forces, with a job but not at work, 6 - Not in labor force
# SCH -- 0 - N/A (less than 3 yrs old), 1 - No, has not attended in the last 3 months, 2 - Yes, public school or public college, 3 - Yes, private school or college or home school

acs1_hh <-
  acs1 %>% 
  .[i = RELSHIPP %in% c(20,21,23), # subset to only householder and spouse records
    by = .(YEAR, SAMPLE, SERIALNO),
    j = .(spouse_present = maxNA(RELSHIPP %in% c(21,23)), # This corresponds to spouse
          all_work_sch   = minNA(ESR %in% 1:2 | SCH %in% 2:3))] # corresponds to working, or in school

  # /!\ Not built, but possibly of interest may be gender of households, and
  # separate examination of SCH and ESR
```


```{r examine build of household fields}
# Allow examination of cases
if (FALSE) {
  check <- 
    merge(acs1[j = .(YEAR, SERIALNO, RELSHIPP, SEX, AGEP, SCH, ESR)],
          acs1_hh[j = .(YEAR, SERIALNO, spouse_present, all_work_sch)],
          all = TRUE) %>% 
    .[order(SERIALNO, YEAR)]
  View(check)
  
  # Calculate how many indicators are missing
  check[j = .(mean(is.na(spouse_present)),
              mean(is.na(all_work_sch)))]
  
}
```

<!-- ### Prepare Child Records -->

```{r view variables relevant to child coding, results = "asis", eval = developer_mode}
if (print_documentation) {
  show_vars <- c("AGEP", "POVPIP", "RAC2P", "HISP", "PWGTP", "YEAR") #, "FAMUNIT" # no variables for famunit 
  my_pums %>% 
    filter(var_code %in% show_vars) %>% 
    kable()
}
```

Besides the indicators above, all other relevant characteristics are measured relative to children themselves (including family income which is reported in records for all members of the household). Development of these data involve:

* Subsetting of children to ages 14 and below
* Inclusion of all measures constructed for heads of household and the household overall
* Age group of child bucketed into intervals of 0-2, 3-5, 6-8, 9-12, and 13-14 <!-- check on the bin_age function which establishes this -->
* Race of child partitioned into "white alone", "Black/African-American alone", and "other"
* Ethnicity of child partitioned into "Hispanic" and "non-Hispanic"
* Family income partitioned into different schemes for use in different modeling specifications, and for alignment to values that CCDF income eligibility thresholds may take on:
  - 0-50% FPL, 50-125%, 125-185%, 185-300%, 300-400%, 400%+ (useful in contexts where the CCDF income eligibility threshold is 185% FPL)
  - 0-50% FPL, 50-100%, 100-200%, 200-300%, 300-400%, 400%+ (useful in contexts where the CCDF income eligibility threshold is 200% FPL)
  - 0-75% FPL, 75-150%, 150-225%, 225-300%, 300-400%, 400%+ (useful in contexts where the CCDF income eligibility threshold is 225% FPL)
  - 0-50% FPL, 50-100%, 100-150%, 150-200%, 200-250%, 250-300%, 300-400%, 400%+ (useful for flexibility despite buckets with relatively few households)
  - 0-100% FPL, 100-200%, 200-300%, 300-400%, 400%+ (useful for simpler reference)
* Combinations of partitions defined above to both represent share categories predicted in the SAE method, as well as direct calculation of CCDF eligibility rates in the baseline year

```{r prepare ACS 1-year microdata for children}
acs1_child <- 
  acs1[AGEP <= 14] %>% 
  merge(acs1_hoh, 
        by = c("YEAR", "SAMPLE", "SERIALNO"), 
        all.x = TRUE) %>%
  merge(acs1_hh,
        by = c("YEAR", "SERIALNO"),
        all.x = TRUE) %>% 
  .[, fam_incpov_ratio := POVPIP / 100] %>% 
  .[, `:=`(child_agegroup = bin_age(AGEP),
           child_race = 
             case_when(RAC2P == 1 ~ "white",
                       RAC2P == 2 ~ "black",
                       TRUE ~ "other_multiple"),
           child_hispanic = 1*(HISP != 1),
           fam_incpov_ratio_cat_mix   = bin_incpov_ratio(fam_incpov_ratio, pov_breaks = c(0, 0.50, 1.00, 2.00, 3.00, 4.00)),
           fam_incpov_ratio_cat_to185 = bin_incpov_ratio(fam_incpov_ratio, pov_breaks = c(0, 0.50, 1.25, 1.85, 3.00, 4.00)),
           fam_incpov_ratio_cat_by50  = bin_incpov_ratio(fam_incpov_ratio, pov_breaks = seq(0, 4, by = 0.5)),
           fam_incpov_ratio_cat_by75  = bin_incpov_ratio(fam_incpov_ratio, pov_breaks = seq(0, 4, by = 0.75)),
           fam_incpov_ratio_cat_by100 = bin_incpov_ratio(fam_incpov_ratio, pov_breaks = seq(0, 4, by = 1.0)),
           obs_weight = PWGTP)] %>% 
  .[j = `:=`(work_spouse_status = paste(ifelse(all_work_sch   == 1, "WorkElig",      "NotWorkElig"),
                                        ifelse(spouse_present == 1, "SpousePresent", "NoSpousePresent"),
                                        sep = "_"),
             work_incpov_status185 = paste(ifelse(all_work_sch == 1, "WorkElig", "NotWorkElig"),
                                        fam_incpov_ratio_cat_to185,
                                        sep = "_"),
             work_incpov_status100 = paste(ifelse(all_work_sch == 1, "WorkElig", "NotWorkElig"),
                                        fam_incpov_ratio_cat_by100,
                                        sep = "_"),
             work_incpov_status75  = paste(ifelse(all_work_sch == 1, "WorkElig", "NotWorkElig"),
                                           fam_incpov_ratio_cat_by75,
                                           sep = "_"))]
```

```{r examine frequencies, eval = developer_mode}
plot_freq <- function(meas) {
  ggplot(acs1_child,
         aes(x = get(meas))) +
    stat_count() + 
    theme(axis.text.x = element_text(angle = 90))
}
plot_freq("fam_incpov_ratio_cat_by50")
plot_freq("fam_incpov_ratio_cat_by100")
plot_freq("work_spouse_status")
plot_freq("work_incpov_status100")
```

```{r verify that data sets only have one obs per child, eval = developer_mode}
acs1_child %>% 
  group_by("SAMPLE", "SERIALNO", "SPORDER") %>% 
  summarize(count = n()) %>% 
  group_by(count) %>% 
  tally() 
```


```{r remove data sets that will not be used in any future data development}
rm(acs1, acs1_hoh)
gc()
```

```{r output basics to be able to compare with direct pulls, eval = developer_mode}
# /*\ This diagnostic was motivated by a specific exercise in building comparisons
# with by-hand pulls of ACS data from ipums.org. If of interest outside of the original
# IL-based application, the income threshold may need modification.
(puma_agg <- 
   acs1_child[between(AGEP, 0, 5),
              .(sample_n = .N,
                n = sum(PWGTP),
                  n_lt225fpl         =   comma(sum(PWGTP[fam_incpov_ratio <= 2.25])),
                pct_lt225fpl         = percent(sum(PWGTP[fam_incpov_ratio <= 2.25]) / sum(PWGTP), accuracy = 0.01),
                  n_lt225fpl_worksch =   comma(sum(PWGTP[fam_incpov_ratio <= 2.25 & all_work_sch == 1])),
                pct_lt225fpl_worksch = percent(sum(PWGTP[fam_incpov_ratio <= 2.25 & all_work_sch == 1]) / sum(PWGTP), accuracy = 0.01)),
              by = PUMA] %>% 
   arrange(PUMA))

if (FALSE) {
  View(puma_agg)
}
```

```{r calculate selected statistics at the PUMA level with standard errors to facilitate later checks}
# Function to calculate poverty rates
calc_acs1_puma_pov <- function(my_data, indices) {
  my_data[indices,
          j = .(pct_pov = meanNA(fam_incpov_ratio <= 1.00)),
          by = .(child_agegroup_combo, PUMA)] %>% 
    .[order(child_agegroup_combo, PUMA)] %>% 
    pull(pct_pov)
}

acs1_child_mod <- 
  acs1_child %>% 
  .[j = child_agegroup_combo := 
      case_when(child_agegroup %in% c("0to2", "3to5") ~ "0to5",
                TRUE ~ as.character(child_agegroup)) %>% 
      factor(levels = c(setdiff(levels(child_agegroup), 
                                c("0to2", "3to5")),
                        "0to5")) %>% 
      fct_relevel("0to5")] %>% 
  .[order(child_agegroup_combo, PUMA)]

acs1_boot <- boot(acs1_child_mod, calc_acs1_puma_pov, R = 200)
acs1_puma_stats <- 
  cbind(acs1_boot$data[j = .(child_agegroup_combo, PUMA)] %>% unique(),
        data.table(pct_pov    = acs1_boot$t0,
                   pct_pov_check = apply(acs1_boot$t, 2, mean)),
                   pct_pov_se = sqrt(apply(acs1_boot$t, 2, var)))
```


```{r}
save(acs1_child,
     acs1_puma_stats,
     file = glue("{output_path}acs1_data_{my_output_tag}.Rda"))
```
